---
title: "matchPreProc"
output: html_document
---

One major problem that occurs in many different tasks is how to reconcile two noisy character based data sets. There are many tools in existence that help facilitate this, but common to all of them is the need to eliminate known sources of noise quickly.

This code preprocesses data prior to use of a series of matching algorithms using string distance and substring operations. It largely achieves this through use of ```grep```, vectorized operations, and parallel operations (i.e. ```sapply```).

An area of optimization yet to occur is the use of distributed processing. With a small enough data set, almost all of the expensive operations could be made across many cores, i.e. using ```paralllel::mcLapply``` or ```parallel::parLapply```. 

Here, the dependencies, environmental variables, and user defined functions are described.
```{r content1, eval=F}
library(RODBC); library(Rcpp);

## Define dir for BLOB xor SQL export
BinaryPath <- FALSE
SqlTable <- FALSE

## Input for reading in data treatment objects
InputPath <- "//tax/"

## UDFs ##

# Remove leading spaces
trim <- function (x) gsub("^\\s+|\\s+$", "", x)

# Given `input' list of space split strings, 
# remove `vector' from all, paste together.
vectorErase <- function(input, vector) {
	subString <- paste0("\\s+",vector,"\\s+", collapse="|")  
	return( 
		paste0(gsub(subString, " ", input, perl=T, ignore.case = T)) 
	)
}

# Retrieve all obs of `vars' from SQL table
GetAllTax <- function(MatchSetChannel,vars) {
	return(
		sqlQuery(MatchSetChannel, 
						 paste0(
						 	"select ",vars," from TaxData "
						 )              
						 ,stringsAsFactors=F)
	)
}

## Compiles, loads appendRcpp()
# Use to preallocate matrix
# Speeds up assign for ragged arrays
# sourceCpp(paste0(InputPath, "cppHelpers.cpp"))
# Deprecated, use if string splitting
```


String split was used previously to analyze and prepare dataset. However, we deemed activity too complicated relative to simple ```grep```. I scripted a function called ```appendRcpp``` if we ever needed to manipulate quantitative data, and wanted to make sure R knew how much memory to apportion for the object.
```{r engine='Rcpp', eval=F}
#include <Rcpp.h>
using namespace Rcpp ;

// [[Rcpp::export]]
void appendRcpp(  List fillVecs, NumericVector newLengths, NumericMatrix retmat, NumericVector retmatLengths ) {
  /* appendRcpp
	Fills numeric matrix
	Loop through rows, filling retmat in with the vectors in list
	then update return matrix size to index the next free
	*/
	
	// Declare vars
	NumericVector fillTmp;
	int sizeOld, sizeNew;
    
	// Pull out dimensions of return matrix to fill
	int nrow = retmat.nrow();
  int ncol = retmat.ncol();
    
	// Check that dimensions match
	if ( nrow != retmatLengths.size() || nrow != fillVecs.size() ) { 
        throw std::range_error("In appendC(): dimension mismatch");
    }
    
	// Traverse ddimensions
	for (int ii=0; ii<ncol; ii++) {
        throw std::range_error("In appendC(): exceeded max cols");
		
	// Iterator for row to fill
        NumericMatrix::Row retRow = retmat(ii, _);
	
  // Fill row of return matrix, starting at first non-zero element
        std::copy( fillTmp.begin(), fillTmp.end(), retRow.begin() + sizeOld );
  
  // Update size of return matrix
        retmatLengths[ii] = sizeNew;
		}
	}
```

Here, all of the previously mentioned functions are called. The overall process takes in the contents of a SQL table, cleans and regularizes variables across the dataset, and then saves to either binary large object (i.e. RData file) or SQL write. 

An additional area of optimization is the ```GetAllTax``` function. Currently, it merely feeds a type of ```SELECT *``` operation to the SQLAgent with no major tuning parameters. Revising the query to only use the smallest necessary set could save a great deal of time. Refactoring the code into a stored procedure call with tuned degree of parallelism would also increase efficiency.
```{r content2, eval=F}
## OPS ##


## Define columns
ColumnNamesMatchandAppend <- read.csv("/ColumnNamesMatchandAppend.csv")
ColumnNamesMatchandAppend$X <- NULL

## Data treatment objects
# For Street Name Suffixes
StreetNamesandAbbreviationsFinal <- read.csv(paste0(InputPath,"/StreetNamesandAbbreviationsFinal.csv"))
StreetNamesandAbbreviationsFinal <- as.character(as.array(StreetNamesandAbbreviationsFinal[,1]))
StreetNamesandAbbreviationsFinal <- toupper(gsub("\\.","\\\\\\.",StreetNamesandAbbreviationsFinal))

# Added to prevent grep compiler crash
StreetNamesandAbbreviationsFinal <- unique(gsub("\\\\\\.","",StreetNamesandAbbreviationsFinal))

# For Street Name Directionals
CompassRoseFinal <- read.csv(paste0(InputPath,"/CompassRoseFinal.csv")) 
CompassRoseFinal <- as.character(as.array(CompassRoseFinal[,1]))

# For Street Number
StreetNumberPossibilities <- read.csv(paste0(InputPath,"/StreetNumberPossibilities.csv"))
StreetNumberPossibilities <- as.character(as.array(StreetNumberPossibilities[,1]))

## For City
TaxDataInformation <- read.csv(paste0(InputPath,"/TaxDataInformation.csv"))
CityDirections <- c("NORTH","SOUTH","EAST","WEST", "NORTHEAST","NORTHWEST","SOUTHEAST","SOUTHWEST")


## Variables for query
vars <- "[FIPS CODE], [ETC] "

## SQL Connection string to data provider
MatchSetChannel <- odbcDriverConnect("driver={SQL Server}; etc.")

## Retrieve data
# Benchmark as needed
# ptm <- proc.time()
MatchSet <- GetAllTax(MatchSetChannel,vars)
# proc.time() - ptm

## Prepare Matchset
# Use only first 5 of zip code
# NOTE: Assume each new vector to consume at least 2 GB of RAM to hold in env at peak
MatchSet$ZipCodeCorrected <- substr(MatchSet$"PROPERTY ZIPCODE",1,5)      

# Use leading 0 for FIPS code
MatchSet$FIPSCorrected <- formatC(MatchSet$"FIPS CODE",width=5,flag = "0")

# Treatment for Street Num
# Benchmark: 5-10 minutes per op, 147 million obs
MatchSet$PropertyHouseNumber1 <- ifelse(is.na(MatchSet$"PROPERTY HOUSE NUMBER PREFIX") | 
																					MatchSet$"PROPERTY HOUSE NUMBER PREFIX"==0 |
																					MatchSet$"PROPERTY HOUSE NUMBER PREFIX" == "","",MatchSet$"PROPERTY HOUSE NUMBER PREFIX")
MatchSet$PropertyHouseNumber2 <- ifelse(is.na(MatchSet$"PROPERTY HOUSE NUMBER") | 
																					MatchSet$"PROPERTY HOUSE NUMBER"==0 |
																					MatchSet$"PROPERTY HOUSE NUMBER" == "","",as.integer(MatchSet$"PROPERTY HOUSE NUMBER"))
MatchSet$PropertyHouseNumber3 <- ifelse(is.na(MatchSet$"PROPERTY HOUSE NUMBER SUFFIX") | 
																					MatchSet$"PROPERTY HOUSE NUMBER SUFFIX"==0 |
																					MatchSet$"PROPERTY HOUSE NUMBER SUFFIX"== "","",MatchSet$"PROPERTY HOUSE NUMBER SUFFIX")
MatchSet$PropertyHouseNumber4 <- ifelse(is.na(MatchSet$"PROPERTY APARTMENT UNIT NUMBER") | 
																					MatchSet$"PROPERTY APARTMENT UNIT NUMBER"==0 |
																					MatchSet$"PROPERTY APARTMENT UNIT NUMBER" == "","",MatchSet$"PROPERTY APARTMENT UNIT NUMBER")
MatchSet$HouseNumberCorrected <- paste0(MatchSet$PropertyHouseNumber1,MatchSet$PropertyHouseNumber2,MatchSet$PropertyHouseNumber3,MatchSet$PropertyHouseNumber4)
MatchSet$HouseNumberCorrected <- gsub("[[:punct:]]| ", "", toupper(as.character(MatchSet$HouseNumberCorrected)))

# Initial treatment for Street Name Directional
MatchSet$DirectionCorrected <- ifelse(is.na(MatchSet$"PROPERTY DIRECTION") | MatchSet$"PROPERTY DIRECTION"==0,"",MatchSet$"PROPERTY DIRECTION")

# Initial treatment for City
MatchSet$CityUpper <- toupper(as.character(MatchSet$"PROPERTY CITY"))
MatchSet$CityCorrected <- ""


# Prepare City using grep function
MatchSet$CityCorrected <- sapply( MatchSet$"PROPERTY CITY", vectorErase, CityDirections )
MatchSet$CityCorrected <- trim(gsub("[[:punct:]]| ", "", toupper(as.character(MatchSet$CityCorrected))))

# Prepare Street Name using grep function.
MatchSet$StreetNameCorrected <- sapply( MatchSet$"PROPERTY STREET NAME", vectorErase, StreetNamesandAbbreviationsFinal )
MatchSet$StreetNameCorrected <- trim(gsub("[[:punct:]]| ", "", toupper(as.character(MatchSet$StreetNameCorrected))))


## Code when we used string split:
# temp <- str_split(z)
# output$z <- MatchSet$StreetNameCorrected <- sapply( 1:length(temp), 
#                                         function(x) vectorErase(temp[[x]], StreetNamesandAbbreviationsFinal) )


# Return for binary
if( BinaryPath & !SqlTable ) save(MatchSet, BinaryPath)

# Return for SQL write
if( !BinaryPath & SqlTable ) sqlSave(MatchSetChannel, MatchSet, SqlTable)

# Write variable error handling
if( (BinaryPath & SqlTable) | (!BinaryPath & !SqlTable) cat("Write failed!\n Select either binary or SQL write.")
```
